\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}

### 2018-02-07 redaction/dhp_july2017.tex

\title{\textsc{Transport costs did decline since 1974:\\Findings from modeling additive costs}\thanks{We are grateful to participants at ETSG 17\textsuperscript{th} meetings in Helsinki, DEGIT XXII in Paris School of Economics, and participants at seminars of the University of California-Irvine, the Universities of Paris-Dauphine Paris-I and Nancy for very useful discussions. Finally, any remaining errors are ours. This paper features an online appendix containing additional results and available on the authors' websites.}}
\author{Guillaume \textsc{Daudin}\thanks{%
Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, 75016 PARIS, FRANCE \newline
Universit\'{e} Paris-Dauphine, PSL Research University, IRD, LEDa, UMR 225, DIAL, 75016 PARIS, FRANCE \newline
Sciences Po, Observatoire Fran\c{c}ais des Conjonctures \'{E}conomiques (OFCE), 75014 PARIS, FRANCE \newline
; email: \url{guillaume.daudin@dauphine.fr}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille1.fr}}\qquad Lise \textsc{Patureau}\thanks{Corresponding author. Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, France;  email: \url{lise.patureau@dauphine.fr} } }

@@

\begin{abstract}
 In contrast to \cite{hummels2007}, we find that composition effects play a minor role in accounting for the reduction of transport costs relative to the value of US imports. Transport cost ``per se'' did decline. This difference of results can be attributed to the new way of modeling the additive component of transport costs we offer. Exploiting information contained in the US imports flows database over 1974-2013, distinguishing by transport mode (air and vessel), three main results emerge. First, the additive component of transport costs is sizeable, with mean values over the period of 1.8\% and 2.9\% of the export price in air and vessel transport respectively. Second, taking additive costs into account significantly improves the fit of the modeling of transport costs. Third, the modeling of the additive component is key to computing  the decomposition of the trend patterns of transport costs, between what is attributable to trade composition effects and changes in the transport costs for given goods and origins.

\end{abstract}

@@

\noindent \emph{JEL classification}: F14, N70, R40 \\
\noindent \emph{Keywords}: Transport costs estimates, non-linear econometrics, period 1974-2013, additive costs

@@

Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split between transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariff  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}. In this vein, \citet{anderson_wincoop_jel} obtain that around 30\% of international trade costs are attributable to transport costs. Equivalently, international transportation costs would represent a 21\% markup over production costs. Further, the sizeable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) suggests a non-negligible impact of transport costs on trade flows. If much of trade policy barriers have been removed over the second half of the 20th century, these findings point out that the transport costs component of the overall trade costs remain large and deserve attention. This is accordingly the focus of the paper.\smallskip

This focus is shared with part of the international trade literature that studies the patterns of trade costs over time, such as \cite{hummels2007} and \cite{Behar_Venables}. Many argue that transport costs have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies (see \citealp{Lafourcade_Thisse}). \cite{Glaeser04} find that, over the twentieth century, the cost of moving goods have declined by over 90\% in real terms. However, \cite{hummels2007} points out that trade composition effects dominate the characterization of time trends of international transport costs and that, paradoxically, transport costs for given goods and origins have not declined much. We show this result is not robust to a more flexible treatment of additive transport costs. In contrast to \cite{hummels2007}, we thus find that composition effects play a minor role in accounting for the trend patterns of transport costs. \medskip

Following \citet{samuelson1954}, standard models of international trade have usually modeled trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis)\footnote{Rigourously speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded. As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as commonly done in the literature.} However, the structure (additive vs ad-valorem) of transport costs is far from being anecdotal. The literature has long emphasized its role in shaping the pattern of trade flows. As pointed out by \cite{alchian}, the relative price of two varieties of some good will depend on the level of trade costs as long as they display an additive component. In this context, the relative demand for more expensive/higher quality product goods increases with trade costs (``shipping the good apples out''). This is consistent with the findings by \citet{hummels_skiba}, who estimate the elasticity of freight rates with respect to price to be well below unity. Also, their estimates imply that doubling freight costs increases average free alongside (fas) export prices by 80 to 141 percent, consistent with high quality goods being sold in markets with high freight costs. \citet{Lashkaripour-mimeo-2017} challenges this view. Highlighting that more expensive goods are systematically heavier (and hence more costly to transport\footnote{Yet, one can be concerned by the generality of this result. While the positive correlation between weight and price seems reasonable for goods from the second industrial revolution like cars, it is dubious in the case of ITC goods which importance has been rising since the end of the 20\textsuperscript{th} century.}), he finds an average transport cost elasticity much closer to 1 using firm-level data on Colombian imports. This result is also confirmed on more aggregate US, sector-level trade data. The debate is not over yet, as recent empirical studies based on micro-level data conversely provide a strong empirical support to the role of additive costs (i.e., cost per unit exported) in international trade costs. Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher unit values for exports to more remote countries, in contradiction with the ad-valorem hypothesis. Recently, \citet{Irrazabal_2015} document on Norwegian firm-level trade data that additive costs represent a sizeable share of international trade costs. Based on US sectoral data, our own estimates also suggest an important part for additive costs in international transport costs.

%\citet{Lashkaripour-mimeo-2017} takes into account the fact
%Yet, one can be concerned by the generality of this result. The positive correlation between weight and price seems reasonable for goods from the second industrial revolution like cars, it is dubious in the case of ITC goods which importance has been rising since 1994 (the end point of Lashkaripour's study). Measurement error in Columbian data??

Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs. \citet{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive trade costs, in addition to the ad-valorem component. A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in ad-valorem costs.\footnote{This is due to the alteration of relative prices in a heterogeneous-firms trade framework. Lower additive trade costs reduce the export market prices of the highly performing exporters relative to lowly performing exporters. This, in turn, shifts relative market shares among exporters towards the highly performing ones - a mechanism that can not occur under ad-valorem costs. Total export sales are then reallocated between a smaller number of high-productivity firms. Therefore, the expected profits from exporting prior to entry increase all other things being equal by more in the case of lower per-unit costs, driving up incentives to enter the industry and bringing eventually the welfare enhancing intra-industry reallocations highlighted in \citet{melitz} framework.} Calibrating on Norwegian firm-level data for 2004, \citet{Irrazabal_2015} find that an additive import tariff reduces welfare and trade by more than an identically-sized ad-valorem tariff. While these results suggest that important welfare gains can be achieved by reducing additive trade costs, not much progress has been done in quantifying such gains. One potential reason is the lack of an empirical characterization of the additive component of trade costs.\footnote{The one exception being \citet{Irrazabal_2015}, upon which we come back later.} One objective of the paper is to fill this gap. \bigskip

Do additive costs matter? To answer this question, we provide an empirical decomposition of the structure of transport costs over time, by explicitly distinguishing between ad-valorem and additive parts. To do so, we exploit information contained in the US Imports flows over 1974-2013, from which we get a measure of international transport costs as the difference between the import and the export prices, on a year-product-partner country basis. Obviously, exploiting this database has drawbacks, in particular as it restricts our scope of study to the monetary dimension of international transport costs. Since our measure is based on the gap between export and import prices, it necessarily omits the other dimension of transport costs related to the time value of goods in transit. According to \citet{anderson_wincoop_jel}, the 21\% markup over production costs coming from transport costs includes both directly measured freight costs (10.7\%) and 9\% tax equivalent of ``indirect'' costs related to the time value of goods on their way to their export market (including holding cost for the goods in transit, inventory cost due to buffering the variability of delivery dates, preparation costs associated with shipment size...). Still, the evidence summarized in \citet{anderson_wincoop_jel} points to a persisting importance of direct transport costs, especially compared to other trade barriers. They remain more important than, e.g., policy barriers (8\% tax equivalent), language barrier (7\%) or information cost (6\%). \citet{Hummels_1999} mentions several papers which all point that transport costs pose a barrier similar in size, or larger than tariffs. In the same vein, \citet{limao_venables} highlight the importance of infrastructures for trade costs in general, through their impact on transport costs. Furthermore, our data is based on a single and reliable customs source with a wide time and geographical coverage, and it allows the computation of a direct transport cost measure. This limits the measurement error bias. Therefore, we strongly believe that exploiting this database may deliver useful and reliable insights on the structure of international trade costs.

By investigating the additive dimension of transport costs, our paper is related to the literature that challenges the dominant role of iceberg costs in international trade.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Alessandria-et-al-AER-2010}, \citet{Hornok-et-al-RES-2015} or \citet{Hornok-et-al-JIE-2015}, which point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} Closely related to our paper is the work by \citet{Irrazabal_2015}, which develop a structural framework for inferring additive trade costs from firm-level trade data. Based on Norwegian exports in 2004, they find that the ratio of additive to ad-valorem trade costs is about 14\% of the export price. Our paper complements their findings in two main respects. First, if our study covers a narrower set of trade costs focused on transport costs, our empirical analysis allows us to provide distinct measures of the levels of both the iceberg and the additive trade costs. Second, we exploit exhaustive information about the imports flows of the US over a large time span from 1974 to 2013. In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time. Despite these differences, our results share with \citet{Irrazabal_2015} the important role of the additive component of international trade costs.\smallskip

Our three main findings can be summarized as follows. First, we show that additive costs are quantitatively sizeable. On average over 1974-2013, the additive cost is estimated to 1.8\% and 2.9\% of the export price in air and vessel transport respectively. This is slightly lower than our estimates for the ad-valorem component (3.5\% and 3.2\% for air/vessel respectively on average over the period). Put it differently, additive costs represent between 40 \% and 50\% of the overall transport costs, depending on the year and the transport model. To the best of our knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs.\footnote{Given their database and their empirical approach, \citet{Irrazabal_2015} can only identify the ratio of the additive to the ad-valorem cost ($\frac{t_{ik}}{\tau_{ik}}$ in our terminology), that they interpret by expressing it in terms of the median export price by country-product ($\bar{\widetilde{p}}_{ik}$). By contrast, our estimation strategy enables us to uncover both values of the ad-valorem and the additive costs $\tau_{ik}$ and $t_{ik}$ separately, in terms of the country-product export price $\widetilde{p}_{ik}$.}. This represents a valuable insight for calibrating related international trade or business cycle models.

Second, we provide an empirical assessment of what standard international trade models lose by skipping additive transport costs. Quantitatively, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor 2. On average over the whole period, biased estimates for ad-valorem are 5\% for air and 6\% for vessel, while unbiased estimates are respectively 2.5\% and 3.2\%. Further, we obtain a systematically better goodness of fit when the additive component is included in the regression, even when taking into account the additional degrees of freedom.

Third, additive costs matter in accounting for the time trend of transport costs, as we show by exploiting the time dimension of our database to document the patterns of transport costs over time. Between 1974 and 2013, we thus find that transport costs have decreased by roughly 50\% in air shipping and 60\% in vessel shipping, and of same order of magnitude for both additive and iceberg components. Importantly, we find that most of this decline can be attributed to a reduction in the ``ceteris paribus'' transport costs, rather than to trade composition effects, in contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004). Importantly, we show that the difference of results can be attributed to the modeling of the additive component: Allowing the additive costs to vary over the triple time/product/country partner case (rather than being constant as in \cite{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period. In all three aspects, our results thus point the importance of the additive component in accounting for international transport costs.\smallskip

>>>>

\title{\textsc{International Transport costs:\\New Findings from modeling additive costs}\thanks{We are grateful to participants at ETSG 17\textsuperscript{th} meetings in Helsinki, DEGIT XXII in Paris School of Economics, and participants at seminars of the University of California-Irvine, the Universities of Paris-Dauphine Paris-I and Nancy for very useful discussions. Finally, any remaining errors are ours. This paper features an online appendix containing additional results and available on the authors' websites.}}
\author{Guillaume \textsc{Daudin}\thanks{%
Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, 75016 PARIS, FRANCE \newline
Universit\'{e} Paris-Dauphine, PSL Research University, IRD, LEDa, UMR 225, DIAL, 75016 PARIS, FRANCE \newline
Sciences Po, Observatoire Fran\c{c}ais des Conjonctures \'{E}conomiques (OFCE), 75014 PARIS, FRANCE \newline
; email: \url{guillaume.daudin@dauphine.fr}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille1.fr}}\qquad Lise \textsc{Patureau}\thanks{Corresponding author. Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, France;  email: \url{lise.patureau@dauphine.fr} } }

@@

\begin{abstract}
 This paper provides a quantitative assessment of the size and the importance of the additive component
in international transport costs, from exploiting information contained in the US imports flows database over 1974-2013 (distinguishing by transport mode, air and vessel). Two main new findings emerge. First, the additive component of transport costs is sizeable, with mean values over the period of 1.8\% and 2.9\% of the export price in air and vessel transport respectively. Second, and in contrast to \cite{hummels2007}, we find that composition effects play a minor role in accounting for the reduction of transport costs relative to the value of US imports. Transport cost ``per se'' did decline. Importantly, this difference of results can be attributed to the new way of modeling the additive component of transport costs we offer. 
\end{abstract}

@@

\noindent \emph{JEL classification}: F14, N70, R40 \\
\noindent \emph{Keywords}: Transport costs estimates, non-linear econometrics, period 1974-2013, additive costs, trade composition effects

@@

Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split between transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariff  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}. In this vein, \citet{anderson_wincoop_jel} obtain that around 30\% of international trade costs are attributable to transport costs. Equivalently, international transportation costs would represent a 21\% markup over production costs. Further, the sizeable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) suggests a non-negligible impact of transport costs on trade flows. If much of trade policy barriers have been removed over the second half of the 20th century, these findings point out that the transport costs component of the overall trade costs remain large and deserve attention. International transport costs accordingly stand at the heart of the paper.\smallskip

This focus is shared with part of the international trade literature that studies the patterns of trade costs over time, such as \cite{hummels2007} and \cite{Behar_Venables}. Many argue that transport costs have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies (see \citealp{Lafourcade_Thisse}). \cite{Glaeser04} find that, over the twentieth century, the cost of moving goods have declined by over 90\% in real terms. However, \cite{hummels2007} points out that trade composition effects dominate the characterization of time trends of international transport costs and that, somehow paradoxically, transport costs ``per se'' (i.e., \textit{ceteris paribus} for given goods and origins) have not declined much. We show this result is not robust to a more flexible treatment of additive transport costs. In contrast to \cite{hummels2007}, we thus find that composition effects play a minor role in accounting for the trend patterns of transport costs, once we allow for a modelling of the additive cost component that varies over time, country partner and product. \medskip

The modelling of transport costs accordingly plays an important role in the paper. Following \citet{samuelson1954}, standard models of international trade have usually modeled trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis).\footnote{Rigourously speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded. As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as commonly done in the literature.} However, in line with the Alchian and Allen's (1964)\nocite{alchian} conjecture, a recent strand of the literature points out to the existence of an additive component (i.e., a cost per unit traded).\footnote{As documented by \cite{Irrazabal_2015}, pricing structure in shipping, additive tariffs, distribution costs... often exhibit (at least partly) an additive structure.} Based on US sectoral data, our own estimates also bring out the importance of additive costs in international transport costs. In this respect, our paper contributes to this literature that challenges the dominant role of iceberg costs in international trade. As pointed out by \cite{alchian}, the relative price of two varieties of some good will depend on the level of trade costs as long as they display an additive component. In this context, the relative demand for more expensive/higher quality product goods increases with trade costs (``shipping the good apples out''). This is consistent with the findings by \citet{hummels_skiba}, who estimate the elasticity of freight rates with respect to price to be well below unity. Also, their estimates imply that doubling freight costs increases average free alongside (fas) export prices by 80 to 141 percent, consistent with high quality goods being sold in markets with high freight costs. \citet{Lashkaripour-mimeo-2017} challenges this view. Highlighting that more expensive goods are systematically heavier (and hence more costly to transport\footnote{Yet, one can be concerned by the generality of this result. While the positive correlation between weight and price seems reasonable for goods from the second industrial revolution like cars, it is dubious in the case of ITC goods whose importance in international trade has been rising since the end of the 20\textsuperscript{th} century.}), he finds an average transport cost elasticity much closer to 1 using firm-level data on Colombian imports. This result is also confirmed on more aggregate US, sector-level trade data. The debate is not over yet, as recent empirical studies based on micro-level data conversely provide a strong empirical support to the role of additive costs (i.e., cost per unit exported) in international trade costs. Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher unit values for exports to more remote countries, in contradiction with the ad-valorem hypothesis.\footnote{Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs. \citet{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive trade costs, in addition to the ad-valorem component. A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in ad-valorem costs, due to the alteration of relative prices in a heterogeneous-firms trade framework. Calibrating on Norwegian firm-level data for 2004, \citet{Irrazabal_2015} find that an additive import tariff reduces welfare and trade by more than an identically-sized ad-valorem tariff. While these results suggest that important welfare gains can be achieved by reducing additive trade costs, not much progress has been done in quantifying such gains. One potential reason is the lack of an empirical characterization of the additive component of trade costs (the one exception being \citet{Irrazabal_2015}). One contribution of the paper is to fill this gap.}
Closely related to our paper is the work by \citet{Irrazabal_2015}, which develop a structural framework for inferring additive trade costs from firm-level trade data. Based on Norwegian firm-level export data in 2004, they find that the ratio of additive to ad-valorem trade costs is about 14\% of the export price. Our paper complements their findings in two main respects. First, if our study covers a narrower set of trade costs focused on transport costs, our empirical analysis allows us to provide distinct measures of the levels of both the iceberg and the additive trade costs. We thus obtain that, on average, the additive cost amounts to 1.8\% and 2.9\% of the export price in air and vessel transport respectively; the average estimated value for the ad-valorem component being of 3.5\% and 3.2\% for air/vessel respectively. Second, we exploit exhaustive information about the imports flows of the US over a large time span from 1974 to 2013. In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time. In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed since 1974.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Alessandria-et-al-AER-2010}, \citet{Hornok-et-al-RES-2015} or \citet{Hornok-et-al-JIE-2015}, which point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip

Do additive costs matter? To answer to this question, we provide an empirical decomposition of the structure of transport costs over time, by explicitly distinguishing between ad-valorem and additive parts. This relies on exploiting information contained in the US Imports flows over 1974-2013, from which we get a measure of international transport costs as the difference between the import and the export prices, on a year-product-partner country basis. Obviously, exploiting this database has drawbacks, in particular as it restricts our scope of study to the monetary dimension of international transport costs. Since our measure is based on the gap between export and import prices, it necessarily omits the other dimension of transport costs notably related to the time value of goods in transit. According to \citet{anderson_wincoop_jel}, the 21\% markup over production costs coming from transport costs includes both directly measured freight costs (10.7\%) and 9\% tax equivalent of ``indirect'' costs related to the time value of goods on their way to their export market (including holding cost for the goods in transit, inventory cost due to buffering the variability of delivery dates, preparation costs associated with shipment size...). Still, the evidence summarized in \citet{anderson_wincoop_jel} points to a persisting importance of direct transport costs, especially compared to other trade barriers. They remain more important than, e.g., policy barriers (8\% tax equivalent), language barrier (7\%) or information cost (6\%). \citet{Hummels_1999} mentions several papers which all point that transport costs pose a barrier similar in size, or larger than tariffs. In the same vein, \citet{limao_venables} highlight the importance of infrastructures for trade costs in general, through their impact on transport costs. Furthermore, our data is based on a single and reliable customs source with a wide time and geographical coverage, and it allows the computation of a direct transport cost measure. This limits the measurement error bias. Therefore, we strongly believe that exploiting this database may deliver useful and reliable insights on the structure of international trade costs.

Our results can be summarized in two main new findings. First, we show that additive costs are quantitatively sizeable. On average over 1974-2013, the additive cost is estimated to 1.8\% and 2.9\% of the export price in air and vessel transport respectively. This is slightly lower than our estimates for the ad-valorem component (3.5\% and 3.2\% for air/vessel respectively on average over the period). Put it differently, additive costs represent between 40 \% and 50\% of the overall transport costs, depending on the year and the transport model. To the best of our knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs.\footnote{Given their database and their empirical approach, \citet{Irrazabal_2015} can only identify the ratio of the additive to the ad-valorem cost ($\frac{t_{ik}}{\tau_{ik}}$ in our terminology), that they interpret by expressing it in terms of the median export price by country-product ($\bar{\widetilde{p}}_{ik}$). By contrast, our estimation strategy enables us to uncover both values of the ad-valorem and the additive costs $\tau_{ik}$ and $t_{ik}$ separately, in terms of the country-product export price $\widetilde{p}_{ik}$.} This represents a valuable insight for calibrating related international trade or business cycle models. We also provide an empirical assessment of what standard international trade models lose by skipping additive transport costs. Quantitatively, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor 2. Further, the goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the additional degrees of freedom.

Second, we establish that the modelling of the additive costs crucially matters in accounting for the trend patterns of international transport costs. Between 1974 and 2013, transport costs have decreased by roughly 50\% in air shipping and 60\% in vessel shipping, and of same order of magnitude for both additive and iceberg components. Importantly, we find that most of this decline can be attributed to a reduction in the ``ceteris paribus'' transport costs, rather than to trade composition effects. This stands in sharp contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004). Importantly, we show that the difference of results can be attributed to the modeling of the additive component: Allowing the additive costs to vary over the triple time/product/country partner case (rather than being constant as in \cite{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period. In both aspects, our results thus point the importance of the additive component in accounting for international transport costs.\smallskip
